AFQ-Insight

### 2019-12-16 reports/manuscript/sections/introduction.tex

To relate the diffusion in each voxel to the macro-structure of
long-range connections between different brain regions, methods for
computational tract-tracing from diffusion MRI, or tractography, combine
the estimates of fiber orientations in each voxel to form streamlines
that traverse the volume of the white matter \cite{Conturo1999-je,
Mori2002-qi}. These methods have been under increased scrutiny and
several lines of investigation have raised critiques of their validity
\cite{Maier-Hein2017-vb, Thomas2014-ki}. On the other hand, there
have been efforts to shore up the inferences made with these methods
\cite{Pestilli2014NatMeth, Takemura2016-sh, Smith2013-nc, Smith2015-cx,
Smith2015-zt, Rheault2018-wk}. Importantly, though discovering novel
tracts requires extraordinary evidence, and delineating
the exact cortical termination of the streamlines in the gray matter
is still prone to error, there is little dispute that tractography
can accurately define the location of several major white matter
tracts that are known to exist within the core of the white matter
\cite{Maier-Hein2017-vb}.

Leveraging this fact, one of the most powerful methods currently
available to put macro- and micro-structure together is
\emph{tractometry}: assessment of the physical properties of the white
matter along specific tracts \cite{Bells2011-cf}. Though there are
several different available implementations of this overall idea,
the principles are similar \cite{yeatman2012tract, Yendiki2011-ay,
Wassermann2016-iv, ODonnell2009-uu}: tractometry begins by delineating
the parts of the white matter that belong to different major ``tracts''
(i.e. anatomical or functional groups of white matter fibers), such as
the corticospinal tract or arcuate fasciculus, assigning tractography
generated streamlines to ``bundles,'' which approximate the anatomical
tracts, and sampling biophysical properties (such as fractional
anisotropy or mean diffusivity) along the length of these bundles.
In some previous tractometry-based studies, tissue properties along
the length of each tract were summarized by taking the mean along
each bundle, but there is a large body of evidence showing that there
is systematic variability along the trajectory of each bundle. This
justifies retaining the individual samples along the length of each
bundle \cite{yeatman2012tract, colby2012, ODonnell2009-uu}. While this
retains important information about each individual's white matter, it
also presents statistical challenges due to the dimensionality of the
data. Based on tractometry, researchers may choose to compare different
individuals to each other. This is usually done according to one of the
following approaches:

>>>>

To relate the diffusion in each voxel to the macro-structure of
long-range connections between different brain regions, methods for
computational tract-tracing from diffusion MRI, or tractography, combine
the estimates of fiber orientations in each voxel to form streamlines
that traverse the volume of the white matter \cite{Conturo1999-je,
Mori2002-qi}. These methods have been under increased scrutiny and
several lines of investigation have raised critiques of their validity
\cite{Maier-Hein2017-vb, Thomas2014-ki}. On the other hand, there
have been efforts to shore up the inferences made with these methods
\cite{Pestilli2014NatMeth, Takemura2016-sh, Smith2013-nc, Smith2015-cx,
Smith2015-zt, Rheault2018-wk}. Importantly, though discovering novel
tracts requires extraordinary evidence, and delineating
the exact cortical termination of the streamlines in the gray matter
is still prone to error, there is little dispute that tractography
can accurately define the location of several major white matter
tracts that are known to exist within the core of the white matter
\cite{Maier-Hein2017-vb, Catani2002-vu}.

Leveraging this fact, one of the most powerful methods currently
available to put macro- and micro-structure together is
\emph{tractometry}: assessment of the physical properties of the white
matter along specific tracts \cite{Bells2011-cf}. Though there are
several different available implementations of this overall idea,
the principles are similar \cite{yeatman2012tract, Yendiki2011-ay,
Wassermann2016-iv, ODonnell2009-uu}: tractometry begins by delineating
the parts of the white matter that belong to different major ``tracts''
(i.e. anatomical or functional groups of white matter fibers), such as
the corticospinal tract or arcuate fasciculus, assigning tractography
generated streamlines to ``bundles,'' which approximate the anatomical
tracts, and sampling biophysical properties (such as fractional
anisotropy or mean diffusivity) along the length of these bundles.
the parts of the white matter that belong to different major tracts
(i.e. anatomical or functional groups of white matter fibers), such as
the corticospinal tract or arcuate fasciculus, assigning tractography
generated streamlines to ``bundles,'' which approximate the anatomical
tracts, and sampling biophysical properties (such as fractional
anisotropy or mean diffusivity) along the length of these bundles. In
some previous tractometry-based studies, tissue properties along the
length of each tract were summarized by taking the mean along each
bundle, but there is a large body of evidence showing that there is
systematic variability in the values of diffusion metrics along the
trajectory of each bundle. This justifies retaining the individual
samples along the length of each bundle \cite{yeatman2012tract,
colby2012, ODonnell2009-uu}. While this retains important information
about each individual's white matter, it also presents statistical
challenges due to the dimensionality of the data. Based on tractometry,
researchers may choose to compare different individuals to each other.
This is usually done according to one of the following approaches:
